---
title: "Machine Learning - Final Project"
author: "MJ"
date: "3 April 2016"
output: word_document
---
```{r, echo=FALSE, results="hide"}
setwd("C:/Users/Michal/Desktop/_data science/08 - Machine Learning/Assignments/Final project/data")

```


```{r, message=FALSE}
library(dplyr)
library(caret)
library(rattle)
library(randomForest)
library(rpart)
```

# Data processing

## Load data

```{r}
training <- read.csv("./data/pml-training.csv", na.strings = c("NA", ""))
testing <- read.csv("./data/pml-testing.csv", na.strings = c("NA", ""))
```

## Edit data

We get rid of columns with missing data.

```{r}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```

We also remove columns predictors that are not relevant to predict the classe (first seven columns)

```{r}
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
```

To be able to get out-of-sample errors we need to split the training data into training and validation sets.

```{r}
set.seed(100) 
inTrain <- createDataPartition(training$classe, p = 0.7, list = FALSE)
train <- training[inTrain, ]
valid <- training[-inTrain, ]
```

#Analysis

We will compare results from two different models - a classifiacation tree and a random forest.

## Classification Tree
```{r}
control <- trainControl(method = "cv", number = 5)
fit_rpart <- train(classe ~ ., data = train, method = "rpart", 
                   trControl = control)
modFitRPART <- train(classe ~ ., method="rpart", data=train, trControl=control)
predRPART <- predict(modFitRPART, valid)
confusionMatrix(predRPART, valid$classe)$overall[1]
```

## Random Forest

```{r}

modFitRF <- train(classe ~ ., data = train, method = "rf",trControl = control)
predRF <- predict(modFitRF, valid)
confusionMatrix(predRF, valid$classe)$overall[1]
```

Random forest appears most accurate and so we use it to predict the test data. The predicted values are then:

```{r}
(predict(modFitRF, testing))
```
